Ensemble Stability of Biological Networks

### 2014-09-23 paper.tex
organization
add definition of graph edit distance and remove unnecessary todo notes

A graph can be decomposed into a natural collection of \emph{modules} called strongly connected components. A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}. The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc} along with their adjacency matrices and hierarchy diagrams (to be defined below).
\todo{add hierarchy diagrams to \ref{fig:scc}}

@@

One can also perform this construction in the oppsite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
a strongly connected graph $C_n$.  To each link $(i,j)$ of $A$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.  The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.
\todo{explain how robustness in \ref{eq:robustness} is symmetric under permutations of the SCCs in $G$ or equivalently of the nodes of $\hier{G}$.}

\todo{add panel to \ref{fig:scc} showing that many graphs get mapped to the same SCC DAG}

\todo{add definition of graph edit distance to the most hierarchical network and define this as our measure of hierarchy}

@@

@@

% \section{Strongly connected components, robustness, and hierarchy}
The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness. Inspection of \ref{eq:sccrobustness} demonstrates that robustness is maximized by maximizing the number of edges \emph{between} the strongly connected components of the graph underlying system interdependencies.
% \section{Hierarchical systems are the most robust}
% \ref{fig:scc} shows systems with varying levels of hierarchy and modularity.

>>>>

A graph can be decomposed into a natural collection of \emph{modules} called strongly connected components. A strongly connected component of a graph is a maximal subset of vertices where each vertex within the subset can be reached from any other \cite{Cormen2009}. The strongly connected components of some examples of three component systems are outlined in \ref{fig:scc} along with their adjacency matrices and hierarchy diagrams (to be defined below).

@@

One can also perform this construction in the opposite direction.  Start with a directed acyclic graph $A$.  To each node $n$ of $A$ associate
a strongly connected graph $C_n$.  To each link $(i,j)$ of $A$ associate a non-empty subset of $\Vertex(C_i) \times \Vertex(C_j)$.  The result will be a graph $G$ such that $\hier(G) = A$ and furthermore, every graph $G$ such that $\hier(G) = A$ can be obtained in this manner.

We note from the definition that the mapping from $G$ to $\hier(G)$ is many-to-one as exemplified in \ref{fig:scc}B and C.
\todo{explain how robustness in \ref{eq:robustness} is symmetric under permutations of the SCCs in $G$ or equivalently of the nodes of $\hier(G)$.}

\todo{add reference for graph edit distance}
The relationship between $G$ and $\hier(G)$ for all $G$ with a given number of vertices suggests a heuristic method of quantifying the degree of hierarchy of a given graph and thus of the system structure it represents. The most hierarchical system is considered to be the graph corresponding to the total ordering, which for three nodes is given in \ref{fig:scc}A. This graph maximizes the number of links between strongly connected components, which also implies maximizing the number of strongly connected components. The graph edit distance on a fixed number of vertices from one graph to another is defined as the minimum number of modifications of the first graph in order to transform it into the second. This distance between any given graph and the total ordering thus quantitatively represents how far a graph is from being perfectly hierarchical. In this work we take this distance to be the definition of hierarchy.

@@

The fact that the most hierarchical of networks is the most robust in the case of three component systems can be understood in general as resulting from the previously mentioned symmetry property of network robustness.

@@

Inspection of \ref{eq:sccrobustness} demonstrates that robustness is maximized by maximizing the number of edges \emph{between} the strongly connected components of the graph underlying system interdependencies.

\todo{write down analogs of \ref{eq:sccrobustness} and \ref{eq:sccmaxrobustness} for resampling a fixed proportion or an arbitrary number of components}

% \section{Strongly connected components, robustness, and hierarchy}

% \section{Hierarchical systems are the most robust}
% \ref{fig:scc} shows systems with varying levels of hierarchy and modularity.
